---
title: "p8105_hw2_xx2485"
author: "Xiaoni Xu"
date: "2024-09-25"
output: github_document
---

Loading needed packages:
```{r, message = FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

Loading and cleaning the NYC transit data; convert the entry variable from character to a logical variable.
```{r} 
transit_df = 
  read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
  
```
This dataset contains the information for every entrance of every subway station of NYC. I chose a number of categories of data from the original dataset, including line, station name, station latitude/longitude, route, entry, vending, entrance type, and ADA compliance. The dimension of the dataset after cleaning is 1868 rows x 19 columns. The data is not quite tidy because `vending` variable is still a YES/NO character variable, the route variable is difficult to understand, and each entrance of the same station is not given a proper way to identify them.

Count distinct stations by both station name and line:
```{r}
distinct_stations <- transit_df %>%
  distinct(station_name, line) %>%
  count()

# View the count of distinct stations
print(distinct_stations)
```
There are 465 distinct stations.

Count ADA-compliant stations
```{r}
ada_compliant_count <- transit_df %>%
  filter(ada == TRUE) %>%
  distinct(station_name, line) %>%
  count()

# View the count
print(ada_compliant_count)
```
There are 84 distinct stations that are ADA compliant.

Calculate the proportion of station entrances / exits without vending allow entrance
```{r}
# Filter for stations without vending and count those allowing entrance
proportion_no_vending_entry <- transit_df %>%
  filter(vending == "NO") %>%    # Select stations with no vending
  summarise(proportion = mean(entry == "TRUE"))  # Calculate the proportion allowing entry

# View the result
print(proportion_no_vending_entry)
```
The proportion of station entrances / exits without vending allow entrance is 0.3770492.


Reformat the data so that route number and route name are distinct
```{r}
transit_df_reformat <- transit_df %>%
  mutate(across(starts_with("route"), as.character)) %>%   # Convert all route columns to character
  pivot_longer(cols = starts_with("route"), # Select all columns that start with 'Route'
               names_to = "route_number",    # Create a new column for the route number
               values_to = "route_name",     # Create a new column for the route name
               values_drop_na = TRUE)        # Drop NA values
```

Find the number of distinct stations that serve the A train
```{r}
# Filter for stations that serve the A train and count distinct stations
distinct_a_train_stations <- transit_df_reformat %>%
  filter(route_name == "A") %>%      
  distinct(station_name, line) %>%     
  count()

# View the count of distinct A train stations
print(distinct_a_train_stations)
```
There are 60 distinct stations that serve the A train.

Find the number of stations that serve the A train and are ADA compliant
```{r}
# Filter for distinct ADA-compliant stations that serve the A train
ada_compliant_a_train_stations <- transit_df_reformat %>%
  filter(route_name == "A", ada == "TRUE") %>%  
  distinct(station_name, line) %>%             
  count()

# View the count of distinct ADA-compliant A train stations
print(ada_compliant_a_train_stations)
```
There are 17 stations that serve the A train and are ADA compliant.


## Problem 2

Read and clean the data
```{r}
# Read and clean the Mr. Trash Wheel data
mr_trash_wheel_df = 
  read_excel("data/202309 Trash Wheel Collection Data.xlsx", 
             sheet = "Mr. Trash Wheel",
             col_names = TRUE) %>% # specify the sheet in the Excel file and to omit non-data entries
  janitor::clean_names() %>% # use reasonable variable names
  select(dumpster:homes_powered)%>% 
  filter(!is.na(dumpster)) %>% # Omit rows without dumpster-specific data
  mutate(sports_balls = as.integer(round(sports_balls))) %>%  # Round sports balls and convert to integer
  mutate(trash_wheel = "Mr. Trash Wheel",
         year = as.integer(year))  # Add identifier

# Read and clean the Professor Trash Wheel data
professor_trash_wheel_df =
  read_excel("data/202309 Trash Wheel Collection Data.xlsx",
                                        sheet = "Professor Trash Wheel",
                                        col_names = TRUE) %>%
  janitor::clean_names() %>%
  select(dumpster:homes_powered) %>%
  filter(!is.na(dumpster)) %>%
  mutate(trash_wheel = "Professor Trash Wheel",
         year = as.integer(year))  # Add identifier

# Read and clean the Gwynnda Wheel data
gwynnda_df <- read_excel("data/202309 Trash Wheel Collection Data.xlsx",
                          sheet = "Gwynnda Trash Wheel",
                          col_names = TRUE) %>%
  janitor::clean_names() %>%
  select(dumpster:homes_powered) %>%
  filter(!is.na(dumpster)) %>%
  mutate(trash_wheel = "Gwynnda",
         year = as.integer(year))  # Add identifier

# Combine all datasets into a single tidy dataset
combined_trash_wheel_df <- bind_rows(mr_trash_wheel_df, 
                                      professor_trash_wheel_df, 
                                      gwynnda_df)

# Display the combined dataset
print(combined_trash_wheel_df)
```
The `combined_trash_wheel_df` dataframe combines the data from three different trash wheels of Healthy Harbor company from the year 2014 to 2023, if data is available. The data includes the number of dumpster, date, weight, volumes, types of trash collected, and homes powered. The combined dataset has 845 rows, with 584 in the Mr. Trash Wheel dataset, 106 in the Professor Trash Wheel dataset, and 155 in the Gwynnda Trash Wheel dataset. 

Calculate the total weight of trash collected by Professor Trash Wheel
```{r}
# Filter for Professor Trash Wheel and pull the weight column
total_weight_professor <- sum(
  combined_trash_wheel_df[combined_trash_wheel_df$trash_wheel == "Professor Trash Wheel", ] %>%
    pull(weight_tons),
  na.rm = TRUE
)

# Display the result
print(total_weight_professor)
```
The total weight of trash collected by Professor Trash Wheel is 216.26 tons.

Calculate the total number of cigarette butts collected by Gwynnda in June of 2022
```{r}
# Filter for Gwynnda in June 2022 and calculate the total cigarette butts
total_cigarette_butts_gwynnda <- combined_trash_wheel_df %>%
  filter(trash_wheel == "Gwynnda" & year == 2022 & month == "June") %>%  
  pull(cigarette_butts) %>%  # Extract the cigarette_butts column
  sum(na.rm = TRUE)  # Calculate the total, ignoring NA values

# Display the result
print(total_cigarette_butts_gwynnda)
```
The total number of cigarette butts collected by Gwynnda in June of 2022 is 18120.




